### 2021-09-09 tex/ms.tex
Rewrite squashing function

\documentclass[a4paper, 12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{dsfont}
\usepackage[left=1.5cm, right=1.5cm, bottom=2cm, top=2cm]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{natbib}
\newcommand{\given}{\,|\,}
\newcommand{\asin}{\sinh^{-1}}

\title{}
\author{Brendon J. Brewer}
\date{}

\begin{document}
\maketitle

%\abstract{}

% Need this after the abstract
\setlength{\parindent}{0pt}
\setlength{\parskip}{8pt}

Let the squashing function be
\begin{align}
S(x) &=
    \left\{
        \begin{array}{lr}
            \log (1 + x), & \textnormal{if } x \geq 0, \\
            -\log(1 - x), & \textnormal{otherwise}.
        \end{array}
    \right.
\end{align}
Operations on the squashed grade can be defined in terms of logsumexp
and logdiffexp, to avoid overflow.

\end{document}