### 2021-10-21 tex/ms.tex
words

\documentclass[a4paper, 12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{dsfont}
\usepackage[left=1.5cm, right=1.5cm, bottom=2cm, top=2cm]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{natbib}
\newcommand{\given}{\,|\,}
\newcommand{\asin}{\sinh^{-1}}

\title{The LBRY Trending Algorithm}
\author{Brendon J. Brewer}
\date{}

\begin{document}
\maketitle

\abstract{I describe the most recent update to the LBRY trending
algorithm. The ideas are similar to previous implementations, but with a
few new features.}

% Need this after the abstract
\setlength{\parindent}{0pt}
\setlength{\parskip}{8pt}

\section{Introduction}
LBRY needs a trending algorithm to rank claims by `recent popularity',
preferably using on-chain information such as changes in support. An
intuitive idea is to have increases and decreases in overall support cause
a `spike' in the trending score, which then decays exponentially over time
(assuming no further changes occur). This has formed the basis of all the
implementations that I have written.

All claims, be they streams/files, channels, reposts, or collections, have
an associated trending score. Hubs (formerly wallet servers) compute and
save the trending score of every claim along with the other data. Several
views of content on LBRY Desktop and Odysee involve sorting claims by their
trending score, in descending order.

In 2021, Jack Robison rewrote a lot of the hub code to improve performance
and to ensure hub data integrity when a chain reorganisation occurs.

\section{Exponentially Decaying Spikes}

\section{Nonlinearity --- the Spike Mass Function}

Let the squashing function be
\begin{align}
S(x) &=
    \left\{
        \begin{array}{lr}
            \log (1 + x), & \textnormal{if } x \geq 0, \\
            -\log(1 - x), & \textnormal{otherwise}.
        \end{array}
    \right.
\end{align}
Operations on the squashed grade can be defined in terms of logsumexp
and logdiffexp, to avoid overflow.

\end{document}